---
title: "Model Comparison"
format: pdf
editor: visual
---

# Setup

```{r}
# install.packages(c("survey", "mgcv", "glmnet", "dplyr", "pROC", "ranger", "xgboost"))

library(survey)
library(mgcv)
library(glmnet)
library(dplyr)
library(pROC)
library(ranger)
library(xgboost)
```

# Data split and survey design

```{r}
df <- readRDS("../data/brfss2024_clean.rds")

predictor_formula <- MEDCOST1 ~
  region +
  rurality +
  sex +
  agegroup +
  race +
  income +
  maritalstatus +
  education +
  employment +
  healthstatus +
  physhlth_days +
  menthlth_days +
  depression +
  chronic_count +
  any_disability +
  insurance +
  pers_doc +
  last_checkup

set.seed(263)
train_index <- unlist(lapply(
  split(seq_len(nrow(df)), df$MEDCOST1),
  function(idx) sample(idx, floor(0.8 * length(idx)))
))

train_df <- df[train_index, ]
test_df <- df[-train_index, ]

y_train <- train_df$MEDCOST1
y_test <- test_df$MEDCOST1

split_summary <- data.frame(
  Set = c("Training", "Testing"),
  N = c(nrow(train_df), nrow(test_df)),
  Event_Rate = round(c(mean(y_train), mean(y_test)), 3)
)
split_summary

# create survey design object for weighted training models
options(survey.lonely.psu = "adjust")

brfss_train_design <- svydesign(
  id = ~PSU,
  strata = ~STSTR,
  weights = ~WEIGHT,
  data = train_df,
  nest = TRUE
)
```

# Evaluation helpers

```{r}
safe_divide <- function(num, den) {
  ifelse(den == 0, NA_real_, num / den)
}

get_auc <- function(y, pred_prob) {
  roc_obj <- roc(y, as.numeric(pred_prob), quiet = TRUE)
  as.numeric(auc(roc_obj))
}

get_youden_threshold <- function(y, pred_prob) {
  roc_obj <- roc(y, as.numeric(pred_prob), quiet = TRUE)
  best <- coords(
    roc_obj,
    x = "best",
    best.method = "youden",
    ret = "threshold"
  )
  as.numeric(best[1])
}

classification_metrics <- function(y, pred_prob, threshold) {
  pred_class <- ifelse(as.numeric(pred_prob) >= threshold, 1, 0)

  tp <- sum(pred_class == 1 & y == 1)
  tn <- sum(pred_class == 0 & y == 0)
  fp <- sum(pred_class == 1 & y == 0)
  fn <- sum(pred_class == 0 & y == 1)

  data.frame(
    Accuracy = mean(pred_class == y),
    Sensitivity = safe_divide(tp, tp + fn),
    Specificity = safe_divide(tn, tn + fp)
  )
}

evaluate_model <- function(model_name, train_y, train_prob, test_y, test_prob) {
  threshold <- get_youden_threshold(train_y, train_prob)
  metrics <- classification_metrics(test_y, test_prob, threshold)

  data.frame(
    Model = model_name,
    Test_AUC = get_auc(test_y, test_prob),
    Threshold = threshold,
    Accuracy = metrics$Accuracy,
    Sensitivity = metrics$Sensitivity,
    Specificity = metrics$Specificity
  )
}
```

# Unweighted models 
## Logistic Regression
```{r}
unweighted_glm <- glm(
  predictor_formula,
  data = train_df,
  family = binomial()
)

summary(unweighted_glm)
```

## Logistic Regression with Lasso
```{r}
x_train <- model.matrix(predictor_formula, data = train_df)[, -1]
x_test <- model.matrix(predictor_formula, data = test_df)[, -1]
```

```{r}
set.seed(263)
unweighted_lasso <- cv.glmnet(
  x_train,
  y_train,
  family = "binomial",
  alpha = 1,
  nfolds = 10,
  type.measure = "auc"
)
```

## GAM
```{r}
unweighted_gam <- gam(
  MEDCOST1 ~
    s(physhlth_days) +
    s(menthlth_days) +
    region +
    rurality +
    sex +
    agegroup +
    race +
    income +
    maritalstatus +
    education +
    employment +
    healthstatus +
    depression +
    chronic_count +
    any_disability +
    insurance +
    pers_doc +
    last_checkup,
  data = train_df,
  family = binomial()
)

summary(unweighted_gam)
```

# Random Forest

```{r}
rf_train_df <- train_df %>%
  mutate(MEDCOST1 = factor(MEDCOST1, levels = c(0, 1)))

rf_test_df <- test_df %>%
  mutate(MEDCOST1 = factor(MEDCOST1, levels = c(0, 1)))

rf_grid <- expand.grid(
  mtry = c(4, 6, 9),
  min.node.size = c(50, 100),
  KEEP.OUT.ATTRS = FALSE
)

set.seed(263)
rf_tuning <- lapply(seq_len(nrow(rf_grid)), function(i) {
  fit <- ranger(
    predictor_formula,
    data = rf_train_df,
    probability = TRUE,
    num.trees = 300,
    mtry = rf_grid$mtry[i],
    min.node.size = rf_grid$min.node.size[i],
    respect.unordered.factors = "order",
    seed = 263
  )

  data.frame(
    mtry = rf_grid$mtry[i],
    min.node.size = rf_grid$min.node.size[i],
    OOB_Error = fit$prediction.error
  )
}) %>%
  bind_rows() %>%
  arrange(OOB_Error)

rf_tuning

best_rf <- rf_tuning[1, ]

set.seed(263)
random_forest <- ranger(
  predictor_formula,
  data = rf_train_df,
  probability = TRUE,
  num.trees = 500,
  mtry = best_rf$mtry,
  min.node.size = best_rf$min.node.size,
  importance = "impurity",
  respect.unordered.factors = "order",
  seed = 263
)

pred_train_rf <- predict(random_forest, data = rf_train_df)$predictions[, "1"]
pred_test_rf <- predict(random_forest, data = rf_test_df)$predictions[, "1"]

rf_importance <- data.frame(
  Variable = names(random_forest$variable.importance),
  Importance = as.numeric(random_forest$variable.importance)
) %>%
  arrange(desc(Importance)) %>%
  mutate(Importance = round(Importance, 3))

head(rf_importance, 15)
```

# Boosting

```{r}
dtrain <- xgb.DMatrix(data = x_train, label = y_train)
dtest <- xgb.DMatrix(data = x_test, label = y_test)

xgb_grid <- expand.grid(
  eta = c(0.05, 0.10),
  max_depth = c(2, 3),
  KEEP.OUT.ATTRS = FALSE
)

scale_pos_weight <- sum(y_train == 0) / sum(y_train == 1)

set.seed(263)
xgb_tuning <- lapply(seq_len(nrow(xgb_grid)), function(i) {
  params <- list(
    objective = "binary:logistic",
    eval_metric = "auc",
    eta = xgb_grid$eta[i],
    max_depth = xgb_grid$max_depth[i],
    min_child_weight = 5,
    subsample = 0.8,
    colsample_bytree = 0.8,
    scale_pos_weight = scale_pos_weight
  )

  cv_fit <- xgb.cv(
    params = params,
    data = dtrain,
    nrounds = 200,
    nfold = 3,
    early_stopping_rounds = 10,
    verbose = 0
  )

  cv_results <- cv_fit$evaluation_log
  best_row <- which.max(cv_results$test_auc_mean)

  data.frame(
    eta = xgb_grid$eta[i],
    max_depth = xgb_grid$max_depth[i],
    best_iteration = cv_results$iter[best_row],
    CV_AUC = cv_results$test_auc_mean[best_row]
  )
}) %>%
  bind_rows() %>%
  arrange(desc(CV_AUC))

xgb_tuning

best_xgb <- xgb_tuning[1, ]

xgb_params <- list(
  objective = "binary:logistic",
  eval_metric = "auc",
  eta = best_xgb$eta,
  max_depth = as.integer(best_xgb$max_depth),
  min_child_weight = 5,
  subsample = 0.8,
  colsample_bytree = 0.8,
  scale_pos_weight = scale_pos_weight
)

set.seed(263)
xgb_model <- xgb.train(
  params = xgb_params,
  data = dtrain,
  nrounds = as.integer(best_xgb$best_iteration),
  verbose = 0
)

pred_train_xgb <- predict(xgb_model, dtrain)
pred_test_xgb <- predict(xgb_model, dtest)

xgb_importance <- xgb.importance(
  feature_names = colnames(x_train),
  model = xgb_model
) %>%
  as.data.frame()

head(xgb_importance, 15)
```


# Weighted models
## Logistic Regression
```{r}
weighted_glm <- svyglm(
  predictor_formula,
  design = brfss_train_design,
  family = quasibinomial()
)

summary(weighted_glm)
```

## Logistic Regression with Lasso
```{r}
set.seed(263)
weighted_lasso <- cv.glmnet(
  x_train,
  y_train,
  family = "binomial",
  alpha = 1,
  nfolds = 10,
  type.measure = "auc",
  weights = train_df$WEIGHT
)
```

## GAM
```{r}
weighted_gam <- gam(
  MEDCOST1 ~
    s(physhlth_days) +
    s(menthlth_days) +
    region +
    rurality +
    sex +
    agegroup +
    race +
    income +
    maritalstatus +
    education +
    employment +
    healthstatus +
    depression +
    chronic_count +
    any_disability +
    insurance +
    pers_doc +
    last_checkup,
  data = train_df,
  family = binomial(),
  weights = train_df$WEIGHT
)

summary(weighted_gam)
```


# Sensitively Analysis
## Logistic Regression
Coefficient stability
```{r}
# model summary comparison table
weighted_glm_summary <- summary(weighted_glm)$coefficients
unweighted_glm_summary <- summary(unweighted_glm)$coefficients

weighted_glm_df <- data.frame(
  Variable = rownames(weighted_glm_summary),
  Weighted_Estimate = weighted_glm_summary[, 1],
  Weighted_SE = weighted_glm_summary[, 2],
  Weighted_p = weighted_glm_summary[, 4]
)

unweighted_glm_df <- data.frame(
  Variable = rownames(unweighted_glm_summary),
  Unweighted_Estimate = unweighted_glm_summary[, 1],
  Unweighted_SE = unweighted_glm_summary[, 2],
  Unweighted_p = unweighted_glm_summary[, 4]
)

# OR and 95% CI
weighted_glm_df$Weighted_OR <- exp(weighted_glm_df$Weighted_Estimate)
weighted_ci <- exp(confint(weighted_glm))
weighted_glm_df$Weighted_OR_lower <- weighted_ci[, 1]
weighted_glm_df$Weighted_OR_upper <- weighted_ci[, 2]

unweighted_glm_df$Unweighted_OR <- exp(unweighted_glm_df$Unweighted_Estimate)
unweighted_ci <- exp(confint.default(unweighted_glm))
unweighted_glm_df$Unweighted_OR_lower <- unweighted_ci[, 1]
unweighted_glm_df$Unweighted_OR_upper <- unweighted_ci[, 2]

comparison_glm_table <- merge(
  weighted_glm_df,
  unweighted_glm_df,
  by = "Variable"
)

comparison_glm_table <- comparison_glm_table %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))

comparison_glm_table <- comparison_glm_table %>%
  mutate(
    Estimate_Diff = Weighted_Estimate - Unweighted_Estimate,
    OR_Diff = Weighted_OR - Unweighted_OR,
    Significant_Weighted = Weighted_p < 0.05,
    Significant_Unweighted = Unweighted_p < 0.05
  )

comparison_glm_table <- comparison_glm_table %>%
  mutate(
    Estimate_Diff = Weighted_Estimate - Unweighted_Estimate,
    OR_Diff = Weighted_OR - Unweighted_OR,
    Direction_Changed =
      sign(Weighted_Estimate) != sign(Unweighted_Estimate),
    Significance_Changed =
      (Weighted_p < 0.05) != (Unweighted_p < 0.05)
  )

glm_sensitivity_summary <- data.frame(
  Metric = c(
    "Number of variables",
    "Variables with direction change",
    "Variables with significance change",
    "Mean absolute coefficient difference",
    "Median absolute coefficient difference",
    "Maximum absolute coefficient difference"
  ),
  Value = c(
    nrow(comparison_glm_table),
    sum(comparison_glm_table$Direction_Changed, na.rm = TRUE),
    sum(comparison_glm_table$Significance_Changed, na.rm = TRUE),
    mean(abs(comparison_glm_table$Estimate_Diff), na.rm = TRUE),
    median(abs(comparison_glm_table$Estimate_Diff), na.rm = TRUE),
    max(abs(comparison_glm_table$Estimate_Diff), na.rm = TRUE)
  )
)

glm_sensitivity_summary

top_changes <- comparison_glm_table %>%
  arrange(desc(abs(Estimate_Diff))) %>%
  select(
    Variable,
    Weighted_OR,
    Unweighted_OR,
    Estimate_Diff,
    Significance_Changed
  ) %>%
  head(10)

top_changes

plot(
  comparison_glm_table$Weighted_OR,
  comparison_glm_table$Unweighted_OR,
  xlab = "Weighted Odds Ratios",
  ylab = "Unweighted Odds Ratios",
  main = "Sensitivity Analysis of Logistic Regression"
)

abline(0, 1, col = "red")

```

Inference and Prediction stability
```{r}
pred_train_w_glm <- as.numeric(predict(weighted_glm, type = "response"))
pred_test_w_glm <- as.numeric(predict(
  weighted_glm,
  newdata = test_df,
  type = "response"
))

pred_train_u_glm <- as.numeric(predict(unweighted_glm, type = "response"))
pred_test_u_glm <- as.numeric(predict(
  unweighted_glm,
  newdata = test_df,
  type = "response"
))

auc_train_w <- auc(train_df$MEDCOST1, pred_train_w_glm)
auc_train_u <- auc(train_df$MEDCOST1, pred_train_u_glm)

auc_test_w <- auc(test_df$MEDCOST1, pred_test_w_glm)
auc_test_u <- auc(test_df$MEDCOST1, pred_test_u_glm)

glm_auc_table <- data.frame(
  Model = c(
    "Weighted Logistic",
    "Unweighted Logistic"
  ),
  Train_AUC = c(
    auc_train_w,
    auc_train_u
  ),
  Test_AUC = c(
    auc_test_w,
    auc_test_u
  )
)

glm_auc_table <- glm_auc_table %>%
  mutate(across(where(is.numeric), round, 3))

glm_auc_table

cor(pred_train_w_glm, pred_train_u_glm)
```


## Logistic Regression with Lasso
Coefficient stability
```{r}
# Lasso selected variables
lasso_out <- cbind(
  coef(weighted_lasso, s = "lambda.1se"),
  coef(unweighted_lasso, s = "lambda.1se")
)
colnames(lasso_out) <- c("Weighted", "Unweighted")
lasso_out

coef_w <- as.matrix(coef(weighted_lasso, s = "lambda.1se"))
coef_u <- as.matrix(coef(unweighted_lasso, s = "lambda.1se"))

weighted_lasso_df <- data.frame(
  Variable = rownames(coef_w),
  Weighted_Estimate = as.numeric(coef_w[, 1])
)

unweighted_lasso_df <- data.frame(
  Variable = rownames(coef_u),
  Unweighted_Estimate = as.numeric(coef_u[, 1])
)

comparison_lasso_table <- merge(
  weighted_lasso_df,
  unweighted_lasso_df,
  by = "Variable",
  all = TRUE
)

# replace NA with 0 (important for LASSO)
comparison_lasso_table[is.na(comparison_lasso_table)] <- 0

comparison_lasso_table <- comparison_lasso_table %>%
  mutate(
    Estimate_Diff = Weighted_Estimate - Unweighted_Estimate,
    
    Direction_Changed =
      (Weighted_Estimate != 0 | Unweighted_Estimate != 0) &
      (sign(Weighted_Estimate) != sign(Unweighted_Estimate)),
    
    Selection_Changed =
      (Weighted_Estimate == 0 & Unweighted_Estimate != 0) |
      (Weighted_Estimate != 0 & Unweighted_Estimate == 0)
  )

lasso_sensitivity_summary <- data.frame(
  Metric = c(
    "Number of variables",
    "Variables with direction change",
    "Variables with selection change",
    "Mean absolute coefficient difference",
    "Median absolute coefficient difference",
    "Maximum absolute coefficient difference"
  ),
  Value = c(
    nrow(comparison_lasso_table),
    sum(comparison_lasso_table$Direction_Changed, na.rm = TRUE),
    sum(comparison_lasso_table$Selection_Changed, na.rm = TRUE),
    mean(abs(comparison_lasso_table$Estimate_Diff), na.rm = TRUE),
    median(abs(comparison_lasso_table$Estimate_Diff), na.rm = TRUE),
    max(abs(comparison_lasso_table$Estimate_Diff), na.rm = TRUE)
  )
)

lasso_sensitivity_summary

top_changes_lasso <- comparison_lasso_table %>%
  arrange(desc(abs(Estimate_Diff))) %>%
  select(
    Variable,
    Estimate_Diff,
    Direction_Changed,
    Selection_Changed
  ) %>%
  head(10)

top_changes_lasso
```

```{r}
pred_train_w_lasso <- as.numeric(predict(
  weighted_lasso,
  newx = x_train,
  s = "lambda.1se",
  type = "response"
))
pred_test_w_lasso <- as.numeric(predict(
  weighted_lasso,
  newx = x_test,
  s = "lambda.1se",
  type = "response"
))

pred_train_u_lasso <- as.numeric(predict(
  unweighted_lasso,
  newx = x_train,
  s = "lambda.1se",
  type = "response"
))
pred_test_u_lasso <- as.numeric(predict(
  unweighted_lasso,
  newx = x_test,
  s = "lambda.1se",
  type = "response"
))
```

Variable selection stability
```{r}
w_coef <- rownames(coef(weighted_lasso, s="lambda.1se"))[
  coef(weighted_lasso, s="lambda.1se")[,1] != 0
]

u_coef <- rownames(coef(unweighted_lasso, s="lambda.1se"))[
  coef(unweighted_lasso, s="lambda.1se")[,1] != 0
]

intersection <- length(intersect(w_coef, u_coef))
union <- length(unique(c(w_coef, u_coef)))

jaccard <- intersection / union
jaccard

coef_diff <- lasso_out[,1] - lasso_out[,2]
summary(coef_diff)

plot(lasso_out[,1], lasso_out[,2],
     xlab="Weighted coef", ylab="Unweighted coef")
abline(0,1)
```


```{r}
par(
  mfrow = c(2, 2),
  mar = c(4, 4, 2, 1),   # reduce top margin
  oma = c(0, 0, 2, 0),   # outer margin for global title
  cex.main = 0.9,       # smaller titles
  cex.axis = 0.8,       # smaller axis labels
  cex.lab = 0.9         # smaller axis titles
)
# Weighted coefficient paths
plot(
  weighted_lasso$glmnet.fit,
  xvar = "lambda",
  label = FALSE,
  ylab = "Weighted LASSO Coefficient Paths"
)

# Weighted CV curve
plot(
  weighted_lasso,
  ylab = "Weighted LASSO Cross-Validation AUC "
)

abline(
  v = log(weighted_lasso$lambda.1se),
  lty = 2
)

# Unweighted coefficient paths
plot(
  unweighted_lasso$glmnet.fit,
  xvar = "lambda",
  label = FALSE,
  ylab = "Unweighted LASSO Coefficient Paths"
)

# Unweighted CV curve
plot(
  unweighted_lasso,
  ylab = "Unweighted LASSO Cross-Validation AUC "
)

abline(
  v = log(unweighted_lasso$lambda.1se),
  lty = 2
)

mtext(
  "Weighted vs Unweighted LASSO Comparison",
  outer = TRUE,
  cex = 1.3
)

par(mfrow = c(1,1))
```


## Comparison

```{r}
# coefficients comparison table
weighted_gam_coef <- summary(weighted_gam)$p.coeff
unweighted_gam_coef <- summary(unweighted_gam)$p.coeff

weighted_gam_df <- data.frame(
  Variable = names(weighted_gam_coef),
  Weighted_Est = weighted_gam_coef,
  Weighted_OR = exp(weighted_gam_coef)
)

unweighted_gam_df <- data.frame(
  Variable = names(unweighted_gam_coef),
  Unweighted_Est = unweighted_gam_coef,
  Unweighted_OR = exp(unweighted_gam_coef)
)

gam_comparison <- merge(weighted_gam_df, unweighted_gam_df, by = "Variable")
gam_comparison <- gam_comparison %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))
gam_comparison

# smooth term plot
par(mfrow = c(2, 2))
plot(weighted_gam, select = 1, main = "physhlth (weighted)")
plot(weighted_gam, select = 2, main = "menthlth (weighted)")
plot(unweighted_gam, select = 1, main = "physhlth (unweighted)")
plot(unweighted_gam, select = 2, main = "menthlth (unweighted)")

# p-values for smooth terms
summary(weighted_gam)$s.pv
summary(unweighted_gam)$s.pv

pred_train_w_gam <- as.numeric(predict(weighted_gam, type = "response"))
pred_test_w_gam <- as.numeric(predict(
  weighted_gam,
  newdata = test_df,
  type = "response"
))

pred_train_u_gam <- as.numeric(predict(unweighted_gam, type = "response"))
pred_test_u_gam <- as.numeric(predict(
  unweighted_gam,
  newdata = test_df,
  type = "response"
))
```


# Tree-based model figures

```{r fig.width=9, fig.height=5}
#png(filename = "../figures/tree_model_variable_importance.png", width = 1200, height = 650, res = 150)

top_rf <- head(rf_importance, 10)
top_xgb <- head(xgb_importance, 10)

old_par <- par(no.readonly = TRUE)
par(mfrow = c(1, 2), mar = c(5, 10, 4, 1))

barplot(
  rev(top_rf$Importance),
  names.arg = rev(top_rf$Variable),
  horiz = TRUE,
  las = 1,
  cex.names = 0.75,
  main = "Random Forest",
  xlab = "Impurity importance",
  col = "#4C78A8",
  border = NA
)

barplot(
  rev(top_xgb$Gain),
  names.arg = rev(top_xgb$Feature),
  horiz = TRUE,
  las = 1,
  cex.names = 0.75,
  main = "XGBoost",
  xlab = "Gain",
  col = "#F58518",
  border = NA
)

par(old_par)
```

```{r fig.width=7, fig.height=5}
#png(filename = "../figures/roc_rf_xgb.png", width = 900, height = 650, res = 150)

roc_rf <- roc(y_test, pred_test_rf, quiet = TRUE)
roc_xgb <- roc(y_test, pred_test_xgb, quiet = TRUE)

plot(
  roc_rf,
  col = "#4C78A8",
  lwd = 2,
  main = "ROC Curves: Random Forest vs XGBoost"
)
plot(
  roc_xgb,
  col = "#F58518",
  lwd = 2,
  add = TRUE
)

legend(
  "bottomright",
  legend = c(
    paste0("Random Forest AUC = ", round(as.numeric(auc(roc_rf)), 3)),
    paste0("XGBoost AUC = ", round(as.numeric(auc(roc_xgb)), 3))
  ),
  col = c("#4C78A8", "#F58518"),
  lwd = 2,
  bty = "n"
)
```

# Final model comparison

```{r}
model_comparison <- rbind(
  evaluate_model(
    "Logistic (Weighted)",
    y_train,
    pred_train_w_glm,
    y_test,
    pred_test_w_glm
  ),
  evaluate_model(
    "Logistic (Unweighted)",
    y_train,
    pred_train_u_glm,
    y_test,
    pred_test_u_glm
  ),
  evaluate_model(
    "LASSO (Weighted)",
    y_train,
    pred_train_w_lasso,
    y_test,
    pred_test_w_lasso
  ),
  evaluate_model(
    "LASSO (Unweighted)",
    y_train,
    pred_train_u_lasso,
    y_test,
    pred_test_u_lasso
  ),
  evaluate_model(
    "GAM (Weighted)",
    y_train,
    pred_train_w_gam,
    y_test,
    pred_test_w_gam
  ),
  evaluate_model(
    "GAM (Unweighted)",
    y_train,
    pred_train_u_gam,
    y_test,
    pred_test_u_gam
  ),
  evaluate_model(
    "Random Forest",
    y_train,
    pred_train_rf,
    y_test,
    pred_test_rf
  ),
  evaluate_model(
    "XGBoost",
    y_train,
    pred_train_xgb,
    y_test,
    pred_test_xgb
  )
)

model_comparison <- model_comparison %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  arrange(desc(Test_AUC))

model_comparison
```

# Visuals

## ROC Curves for Five Unweighted Models

```{r}
library(pROC)

# ROC objects for five unweighted / non-survey-weighted models
roc_logistic <- roc(y_test, pred_test_u_glm, quiet = TRUE)
roc_lasso    <- roc(y_test, pred_test_u_lasso, quiet = TRUE)
roc_gam      <- roc(y_test, pred_test_u_gam, quiet = TRUE)
roc_rf       <- roc(y_test, pred_test_rf, quiet = TRUE)
roc_xgb      <- roc(y_test, pred_test_xgb, quiet = TRUE)

roc_list <- list(
  "Logistic Regression" = roc_logistic,
  "LASSO" = roc_lasso,
  "GAM" = roc_gam,
  "Random Forest" = roc_rf,
  "XGBoost" = roc_xgb
)

cols <- c("#1B9E77", "#D95F02", "#7570B3", "#4C78A8", "#F58518")

# Save ROC curve as PNG
#png("../figures/roc_curves_5_models.png", width = 900, height = 700, res = 120)

plot(
  roc_list[[1]],
  col = cols[1],
  lwd = 2,
  legacy.axes = TRUE,
  main = "ROC Curves for Five Models",
  xlab = "1 - Specificity",
  ylab = "Sensitivity"
)

for (i in 2:length(roc_list)) {
  plot(
    roc_list[[i]],
    col = cols[i],
    lwd = 2,
    add = TRUE
  )
}

legend(
  "bottomright",
  legend = paste0(
    names(roc_list),
    " AUC = ",
    round(sapply(roc_list, function(x) as.numeric(auc(x))), 3)
  ),
  col = cols,
  lwd = 2,
  cex = 0.8,
  bty = "n"
)

#dev.off()
```

# Table of five unweighted models

```{r}
library(dplyr)
library(knitr)

table_5_unweighted <- model_comparison %>%
  filter(Model %in% c(
    "Logistic (Unweighted)",
    "LASSO (Unweighted)",
    "GAM (Unweighted)",
    "Random Forest",
    "XGBoost"
  )) %>%
  arrange(desc(Test_AUC))

kable(
  table_5_unweighted,
  digits = 3,
  align = "lrrrrr"
)
```

# Table of weighted versus unweighted comparison for regression-based models

```{r}
table_weighted_vs_unweighted <- model_comparison %>%
  filter(Model %in% c(
    "Logistic (Weighted)",
    "Logistic (Unweighted)",
    "LASSO (Weighted)",
    "LASSO (Unweighted)",
    "GAM (Weighted)",
    "GAM (Unweighted)"
  )) %>%
  mutate(
    Base_Model = case_when(
      grepl("Logistic", Model) ~ "Logistic Regression",
      grepl("LASSO", Model) ~ "LASSO",
      grepl("GAM", Model) ~ "GAM"
    ),
    Weighting = case_when(
      grepl("\\(Weighted\\)", Model) ~ "Weighted",
      grepl("\\(Unweighted\\)", Model) ~ "Unweighted"
    )
  ) %>%
  select(
    Base_Model,
    Weighting,
    Test_AUC,
    Threshold,
    Accuracy,
    Sensitivity,
    Specificity
  ) %>%
  arrange(Base_Model, Weighting)

kable(
  table_weighted_vs_unweighted,
  digits = 3,
  align = "llrrrrr"
)
```
